Enforced Invariant Confluence

### 2020-08-07 doc/the_vldb_journal_paper/rebuttal.tex

\paragraph{Reviewer 3: Reviewer 3: Other Comments}
\begin{feedback}
  Figure 5 should appear before Figure 6.
\end{feedback}
\textcolor{red}{TODO.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer 4}
\paragraph{Reviewer 4: Overview}
\begin{feedback}
  The paper is very well written, and it makes substantial improvement over the
  VLDB 2019 conference version.
\end{feedback}
Thanks! We appreciate it!

\paragraph{Reviewer 4: 1}
\begin{feedback}
  In the introduction the concept of "invariant" is missing some context. Where
  does this invariant come from? It looks like this invariant is a given
  invariant for the program executing on the replicated object, and it is not
  necessarily a tight invariant.  Clarifying this in advance would be useful,
  as I realized this only in the beginning of Section 2, I read this:
  "Informally, a replicated object is invariant confluent *with respect to an
  invariant* if every replica of the object is guaranteed to satisfy the
  invariant despite the possibility of different replicas being concurrently
  modified or merged together." Mentioning the system model explicitly in the
  introduction also would help: "a distributed object and accompanying
  invariant is replicated across a set of servers."
\end{feedback}

\paragraph{Reviewer 4: 2}
\begin{feedback}
  Invariant confluence is about safety, however, could there be a liveness part
  of the problem worth investigating/discussing for replicated object
  management? Would it be possible that forgoing coordination across replicas
  could jeopardize/violate some liveness properties?
\end{feedback}
Forgoing coordination across replicas in favor of \invariantconfluence{}
actually \emph{increases} liveness, rather than jeopardizing or violating it.
Any alive replica of an \invariantconfluent{} object can service client
requests and execute transactions. This is the strongest form of liveness we
can hope to achieve. In fact, if we aim for strongly consistent replication,
then the CAP theorem tells us that total availability (the kind that
\invariantconfluence{} achieves) is impossible~\cite{brewer2012cap,
gilbert2002brewer}.

Note, however, that a segmented \invariantconfluent{} object does suffer from
liveness problems, as we discuss in Section 6.4. Specifically, transactions
that require global coordination are live only as long as every replica is
live. We discuss this more in Reviewer 4: 5 below.

@@

\paragraph{Reviewer 4: 3}
\begin{feedback}
  Mergeable Replicated Data Types paper (OOPSLA'19) is relevant work. The paper
  does not cite MRDT understandably because it is very recent work, however
  including a discussion of this in the related work would be helpful.
\end{feedback}
\textcolor{red}{TODO.}

\paragraph{Reviewer 4: 4}
\begin{feedback}
  Section 4 is about invariant tightening, which has a broader applicability in
  distributed systems/programs. Unfortunately the techniques used here are not
  new. Involving the user for identifying unreachable states in not practical,
  and the discussion on tolerating user errors is weak as it does not offer a
  remedy. Would using some heuristics and abstracting state provide any help
  here?
\end{feedback}
\textcolor{red}{TODO.}

\paragraph{Reviewer 4: 5}
\begin{feedback}
  Section 6 on segmented invariant confluence is nice. It would be nice if
  Section 6.4 could be improved. I liked the fourth point in Section 6.4, would
  it be possible to elaborate on this. The sixth point however is bad, as
  replicating every server with state-machine-replication is overkill. Would it
  be possible to use use a membership service approach (as in Vertical Paxos
  and Chain Replication configuration) to prevent a single crash to flunk
  global coordination?
\end{feedback}
We added a couple paragraphs to Example 13 with a concrete example of the
optimization described in our fourth point. We also added a forward reference
in Section 6.4 to Example 13.

Your comment on our sixth point is interesting. We hadn't thought of that
before. Coincidentally, since our initial publication, we spent some time
researching membership changes in state machine replication protocols---which
you can find here:
\url{https://mwhittaker.github.io/publications/matchmaker_paxos.html}---so we
can answer your question with a bit of confidence. Assume a client sends a
transactions to a server $p$. The transaction happens to not need global
coordination, so $p$ executes the transaction locally and sends back the result
to the client. This scenario makes membership changes difficult for the
following reason. Consider a different execution in which $p$ is partitioned
from the rest of the servers. The system erroneously thinks that $p$ has failed
and reconfigures to some other set of servers. However, $p$ and the client are
oblivious to this. The client can still contact $p$ and have transactions
executed, even though the system has moved to a completely new set of servers.
This can lead to situations that are impossible without membership changes. For
example, $p$ may be in a different active segment than the rest of the servers.

This problem can be mitigated by making stronger assumptions (e.g., a failure
detector, clock synchrony assumptions with leases). Alternatively, we could
require that a client contact multiple servers to execute a transaction and
then use quorum intersection to perform membership changes safely. We think
that this is a very interesting direction of future work, but unfortunately (as
with all membership change protocols) is very nuanced and very easy to get
wrong. It is out of the scope of our paper.

\begin{feedback}
  Is there plans to make Lucy implementation open source?
\end{feedback}
Yes. Lucy is available online at
\url{https://github.com/mwhittaker/enforced_invariant_confluence}. We also
added this link to Section 8 of the paper.

\paragraph{Reviewer 4: 7}
\begin{feedback}
  In the throughput evaluation of segmented invariant confluence systems, what
  would be the effects/overhead of faulttolerance? (Related to item 5 above.)
\end{feedback}
If we replicate every server with a state machine replication protocol, the
throughput of the segmented \invariantconfluent{} object certainly
decreases. However, since ever server can still independently execute
transactions, the throughput of the system would still be scalable for
workloads that don't require frequent coordination.
%
The performance overheads of more complex forms of fault tolerance (as
described in item 5 above) are out of scope for our paper.

\paragraph{Reviewer 4: 8}
\begin{feedback}
  Would it be possible to elaborate more on escrow transaction implementation.
  It was not clear to me from Example 13.
\end{feedback}
We added two more paragraphs of explanation to Example 13, including a concrete
example. Hopefully, that makes things clearer.

### 2020-08-07 doc/the_vldb_journal_paper/sections/evaluation.tex
Start on revisions.
More progress on rebuttal.
More work on revisions.

\section{Evaluation}\seclabel{Evaluation}
In this section, we describe and evaluate Lucy: a prototype implementation of
our decision procedures and system models.

@@

\example[Auction]\examplelabel{AuctionEval}
We now consider a simple auction system introduced in~\cite{gotsman2016cause}.
Our object consists of a set $B$ of integer-valued bids and a optional winning
bid $w$. Initially, $B = \emptyset$ and $w = \bot$ (indicating that there is no
winning bid yet) and we merge states by taking the union of $B$ and the maximum
of $w$ (where $\bot < n$ for all integers $n$). One transaction $t_b$ places a
bid $b$ by inserting it into $B$. Another transaction $t_\text{close}$ closes
the auction and sets $w$ equal to the largest bid in $B$. Our invariant is that
if the auction is closed (i.e.\ $w \neq \bot$), then $w = \max(B)$. We ran our
decision procedure on this example and in a third of a second, it produced a
reachable counterexample witnessing the fact that the object is not
\invariantconfluent{}.  If we concurrently close the auction and place a large
bid, then we can end up in a state in which the auction is closed, but there is
a bid in $B$ larger than $w$.

@@

Consider again, from \exampleref{CounreachableExample}, the PN-Counter $s =
(p_1, p_2, p_3)$, $(n_1, n_2, n_3)$ replicated on three servers with
transactions to increment and decrement the PN-Counter. In
\exampleref{CounreachableExample}, we found that concurrent decrements violate
\invariantconfluence{} which led us to a segmentation which prohibited
concurrent decrements. We now propose a new segmentation with escrow amount $k$
that will allow us to perform concurrent decrements that respect the escrowed
value. The first segment is $(I_k, T)$ where
\[
  I_k = \setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}{n_1, n_2, n_3 \leq k \leq p_1, p_2, p_3}
\]
This segment allows for concurrent increments and decrements so long as every
$p_i \geq k$ and every $n_i \leq k$.  Intuitively, this segment represents the
situation in which every server has escrowed a value of $k$. They can decrement
freely, so long as they don't exceed their escrow budget of $k$. The second
segment is the one presented in \exampleref{CounreachableExample} which
prohibits concurrent decrements. We ran our decision procedure on this example
and it concluded that it was segmented \invariantconfluent{} in less than a
tenth of a second and without any human interaction.

>>>>

\section{Evaluation}\seclabel{Evaluation}
In this section, we describe and evaluate Lucy: a prototype implementation of
our decision procedures and system models.
\begin{revisions}
  All code is available online:
  \urlstyle{rm}
  \url{https://github.com/mwhittaker/enforced_invariant_confluence}.
\end{revisions}

@@

\example[Auction]\examplelabel{AuctionEval}
We now consider a simple auction system introduced in~\cite{gotsman2016cause}.
Our object consists of a set $B$ of integer-valued bids and an optional winning
bid $w$. Initially, $B = \emptyset$ and $w = \bot$ (indicating that there is no
winning bid yet) and we merge states by taking the union of $B$ and the maximum
of $w$ (where $\bot < n$ for all integers $n$). One transaction $t_b$ places a
bid $b$ by inserting it into $B$. Another transaction $t_\text{close}$ closes
the auction and sets $w$ equal to the largest bid in $B$. Our invariant is that
if the auction is closed (i.e.\ $w \neq \bot$), then $w = \max(B)$. We ran our
decision procedure on this example and in a third of a second, it produced a
reachable counterexample witnessing the fact that the object is not
\invariantconfluent{}.  If we concurrently close the auction and place a large
bid, then we can end up in a state in which the auction is closed, but there is
a bid in $B$ larger than $w$.

@@

Consider again, from \exampleref{CounreachableExample}, the PN-Counter $s =
(p_1, p_2, p_3)$, $(n_1, n_2, n_3)$ replicated on three servers with
transactions to increment and decrement the PN-Counter. In
\exampleref{CounreachableExample}, we found that concurrent decrements violate
\invariantconfluence{} which led us to a segmentation which prohibited
concurrent decrements. We now propose a new segmentation with escrow amount $k$
that will allow us to perform concurrent decrements that respect the escrowed
value. The first segment is $(I_k, T)$ where
\[
  I_k = \setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}%
              {n_1, n_2, n_3 \leq k \leq p_1, p_2, p_3}
\]
This segment allows for concurrent increments and decrements so long as every
$p_i \geq k$ and every $n_i \leq k$.  Intuitively, this segment represents the
situation in which every server has escrowed a value of $k$. They can decrement
freely, so long as they don't exceed their escrow budget of $k$. The second
segment is the one presented in \exampleref{CounreachableExample} which
prohibits concurrent decrements.

\begin{revisions}
  For example, assume that $k = 3$ and that the three states are $s_1 = s_2 =
  s_3 = (p_1, p_2, p_3), (n_1, n_2, n_3) = (3, 4, 5), (3, 3, 3)$. Here, the
  value of the PN-counter is $(3 + 4 + 5) - (3 + 3 + 3) = 3$, which is
  non-negative as expected. If server 1 receives a decrement request, it
  \emph{cannot} execute the decrement without global coordination because doing
  so would cause $n_1$ to exceed the escrow threshold $k = 3$. Naively, it
  seems safe. Server 1 would transition to state $(3, 3, 5), (3, 3, 3)$ which
  has non-negative value $2$. However, if servers 2 and 3 concurrently perform
  decrements of their own, the three servers' states would merge to a negative
  value. Thus, the decrement requires global coordination to rule out the
  possibility of concurrent decrements.

  On the other hand, if server 2 receives a decrement request, it can safely
  execute it locally and transition to state $(3, 3, 5), (3, 3, 3)$. This is
  safe because server 2 knows that every server $i$ maintains $p_i \geq n_i$.
  Thus, after merging, the value of the PN-counter is $\sum_i p_i - \sum_i n_i
  = \sum_i (p_i - n_i)$ which is the sum of non-negative terms. Intuitively, if
  every server agrees to never go in debt, then the PN-counter in aggregate
  never goes in debt.
\end{revisions}

We ran our decision procedure on this example and it concluded that it was
segmented \invariantconfluent{} in less than a tenth of a second and without
any human interaction.

\begin{revisions}
  In \secref{SegmentedDiscussion}, we discussed an optimization in which a
  server receives a transaction $t$ and concludes that the transaction requires
  global coordination. The server can defer the execution of $t$,
  executing other transactions in the mean time. When the server tries to
  execute $t$ a second time, the transaction may no longer require global
  coordination.

  Escrow transactions are a good example where this optimization works
  particularly well. Assume that $k = 3$ and that the three states are $s_1 =
  s_2 = s_3 = (p_1, p_2, p_3), (n_1, n_2, n_3) = (3, 3, 3), (3, 3, 3)$. Here,
  no server can process a decrement transaction locally. If a server receives a
  decrement transaction, it must execute it with global coordination. However,
  if the server instead defers the execution of the decrement transaction and
  happens to receive a number of increment transactions, then it can execute
  the decrement locally, without the need for global coordination.
\end{revisions}

### 2020-08-07 doc/the_vldb_journal_paper/sections/segmented_invariant_confluence.tex
More work on revisions.

\subsection{Discussion and Limitations}
There are a few things worth noting about segmented \invariantconfluence{}, its
system model, and its decision procedure.
%
First, \invariantconfluence{} is a very coarse-grained property. If an object
is \invariantconfluent{}, then we can replicate it with no coordination. If it
is not \invariantconfluent{}, then we have no guarantees. There's no
in-between.  Segmented \invariantconfluence{}, on the other hand, is a much
more fine-grained property that can be applied to applications with varying
degrees of complexity. Segmented \invariantconfluence{} provides guarantees to
complex applications that require a large number of segments and to simple
applications with a smaller number of segments, whereas \invariantconfluence{}
only provides guarantees to applications that can be segmented into a single
segment.

@@

Fourth, we can integrate a couple of optimizations into our system model to
further reduce the amount of coordination it requires. To begin, if a server
with state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t
\in T_i$ to execute, and $t(s_i)$ violates the active invariant but not the
global invariant, instead of initiating a round of global coordination, $p_i$
can simply buffer $t$ for re-execution at a later time. While this increases
the latency required to execute $t$, it's possible that after other
transactions are executed, re-executing $t$ may lead to a state that either
satisfies the active invariant or violates the global invariant. In either
case, a round of global coordination is avoided. Similarly, servers can
buffer transactions that require global coordination, executing an entire
batch of these transactions during a single round of global coordination.

>>>>

\subsection{Discussion and Limitations}~\seclabel{SegmentedDiscussion}
There are a few things worth noting about segmented \invariantconfluence{}, its
system model, and its decision procedure.
%
First, \invariantconfluence{} is a very coarse-grained property. If an object
is \invariantconfluent{}, then we can replicate it with no coordination. If it
is not \invariantconfluent{}, then we have no guarantees. There's no
in-between.  Segmented \invariantconfluence{}, on the other hand, is a much
more fine-grained property that can be applied to applications with varying
degrees of complexity. Segmented \invariantconfluence{} provides guarantees to
complex applications that require a large number of segments and to simple
applications with a smaller number of segments, whereas \invariantconfluence{}
only provides guarantees to applications that can be segmented into a single
segment.

@@

Fourth, we can integrate a couple of optimizations into our system model to
further reduce the amount of coordination it requires. To begin, if a server
with state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t
\in T_i$ to execute, and $t(s_i)$ violates the active invariant but not the
global invariant, instead of initiating a round of global coordination, $p_i$
can simply buffer $t$ for re-execution at a later time. While this increases
the latency required to execute $t$, it's possible that after other
transactions are executed, re-executing $t$ may lead to a state that either
satisfies the active invariant or violates the global invariant. In either
case, a round of global coordination is avoided. \markrevisions{Later, in
\exampleref{EscrowTransactionsEval}, we'll see a concrete example of this
optimization.}
%
Similarly, servers can buffer transactions that require global coordination,
executing an entire batch of these transactions during a single round of global
coordination.
